---
title: "Car Transmission vs. Fuel Consumption -- Regression Models"
author: "fjelltronen"
date: "May 13, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

This report explores the relationship between a set of car-related variable and the miles-per-gallon attribute. First, we introduce the `mtcars` dataset. Then we fit a linear model and use its coeficients to determine that manual transmissions lead to a better fuel consumption when compared with automatic ones, by an estimated 7.245 miles-per-gallon.

### MTCARS Dataset

The Motor Trend Car Road Tests data consists of fuel consumption and 10 aspects of automobile design/performance for 32 cars. A brief analysis of the data is shown in the Appendix below.

```{r, echo=FALSE}
data(mtcars)
```

Convert some of the numeric variables into factors, including `am` (transmission information, where `0` is automatic and `1` is manual):

```{r}
mtcars$am <- factor(mtcars$am)
```

A quick exploration of the relationship between the `am` and `mpg` variables is shown graphically in the Appendix.

## Is an automatic or manual transmission better for `mpg`? / Quantify the `mpg` difference between automatic and manual transmissions

Let us fit a linear model with `mpg` as the outcome and `am` as the regressor. The coeficients of this least square regression model can help answer these questions. 

```{r}
fit <- lm(mpg~am, data = mtcars)
coefficients(fit)
```

The adjusted estimate for the expected change in `mpg` when comparing automatic transmission (`(Intercept)` or `am0`) and manual transmission (`am1`) is `7.245`. This positive number indicates that the automatic transmission is worse for `mpg` (less miles-per-gallon imply worse fuel consumption). More specifically, the estimated difference in `mpg` between automatic and manual transmissions is 7.245 (the mean of `am1` is `24.392`, `7.245` higher than the mean of `am0`, which is `17.147`).

```{r}
summary(fit)
```

Since the p-value of this coefficient is `0.000285` (< 0.05), the t-test for H_0: no difference between `am0` and `am1` versus H_a: difference between `am0` and `am1` is *significant*.

Furthermore, in order to evaluate the fit and residuals of the generated model, we use diagnostic plots.

```{r}
par(mfrow = c(2, 2)); plot(fit)
```

The model does not capture the variance in `mpg` for each type of transmission. Its residuals are normally distributed and spred equally along the range of predictors. No cases are influential with respect to Cook's distance.

## Appendix

### `mtcars` dataset

```{r}
# structure of the data set
str(mtcars)
# value counts for `am`
table(mtcars$am)
# mpg for automatic/manual transmission
library(ggplot2)
library(gridExtra)
g1 <- ggplot(data = mtcars, aes(x = am, y = mpg, color = am)) + 
  geom_boxplot() + geom_point(size = 3, alpha = 0.4) + 
  labs(title = "mpg vs. am", 
       x = "transmission (0 = automatic, 1 = manual)", y = "fuel consumption")
g2 <- ggplot(data = mtcars, aes(x = mpg, color = am)) + geom_density() +
  labs(title = "mpg density")
grid.arrange(g1, g2, ncol = 2)
```



