---
title: "Regression Models -- Analysis of the MTCARS Dataset"
author: "fjelltronen"
date: "May 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

This report explores the relationship between a set of car-related variable and the miles-per-gallon attribute. First, we introduce the `mtcars` dataset. Then we ... 

### MTCARS Dataset

The Motor Trend Car Road Tests data consists of fuel consumption and 10 aspects of automobile design/performance for 32 cars. A brief analysis of the data is shown in the Appendix below.

```{r, echo=FALSE}
data(mtcars)
```

Convert some of the numeric variables into factors, including `am` (transmission information, where `0` is automatic and `1` is manual):

```{r}
mtcars$am <- factor(mtcars$am)
```

A quick exploration of the relationship between `am` and `mpg` is shown below.

```{r}
library(ggplot2)
ggplot(data = mtcars, aes(x = am, y = mpg)) + geom_point(size = 3, alpha = 0.4, color = "lightseagreen") +
  labs(title = "mpg for automatic/manual transmissions", x = "transmission", y = "fuel consumption")
```

## Is an automatic or manual transmission better for MPG?

## Quantify the MPG difference between automatic and manual transmissions

## Appendix

### `mtcars` dataset

```{r}
# structure of the data set
str(mtcars)
```
